\large{\bf Looking forwards and backwards: dynamics and genealogies of locally regulated populations }

### 2023-04-25 forwards-and-backwards.tex
pass through remaining things

@@

\paragraph{Scaling limits with nonlocal interactions:}
Recall that the process $(\eta_t^N)_{t\geq 0}$ takes its values in the space
${\cal D}([0,\infty),\measures)$ of c\`adl\`ag
paths on $\measures$. We endow $\measures$ with the topology of weak convergence
and 
${\cal D}([0,\infty),\measures)$ with the Skorohod topology.
A sequence of processes taking values in 
${\cal D}([0,\infty),\measures)$ is said to be tight if the corresponding
sequence of distributions is tight,
i.e., that any infinite subsequence has a weakly convergent subsubsequence.
Our first main result establishes tightness of our rescaled population processes in 
the case in which interactions remain nonlocal under the scaling, and 
characterises limit points as solutions to a martingale problem.

@@

\begin{theorem}[Convergence to a PDE]
    \label{thm:local_convergence}
    Let $(\eta^N_t)_{t \geq 0}$
    be as defined in Definition~\ref{defn:mgale_construction} with
$r(x,m)\equiv 1\equiv \gamma(r,m)$, $F(x,m)\equiv F(m)$, 
$\rho^\epsilon_F$ a symmetric Gaussian density with variance
parameter $\epsilon^2$,
and $\DG=\Delta$. 
Further suppose that $F(m)$ is a polynomial with $F(m)\ind_{m\geq 0}$ bounded above.
Assume that $\langle 1,\eta_0\rangle$ is bounded, and that for all 
$x\in\IR^d$ and $k\in \IN$,
	\[
		\limsup_{\epsilon\to 0} 
		\IE\big[\rho_F^\epsilon*\eta_0(x)^k\big]<\infty,
	\]
	and 
\[
    \limsup_{\epsilon \to 0} \int \IE\big[\rho_F^\epsilon*\eta_0(x)^k\big] dx<\infty.
\]
Finally assume that $N \to \infty$, $\theta \to \infty$
and $\epsilon \to 0$ in such a way that
\begin{equation}
\frac{1}{\theta\epsilon^2}+\frac{\theta}{N\epsilon^d}\to 0.
\end{equation}
Then the sequence of 
${\cal D}([0,\infty), \measures)$-valued
stochastic processes $\big(\rho_F^\epsilon*\eta^N_t(x)dx\big)_{t \ge 0}$
converges weakly
to a measure-valued process with a density $\varphi(t, x)$
that solves
\begin{align}
	\label{target equation in local convergence}
        \partial_t \varphi(t, x)=\Delta \varphi (t,x) +\varphi(t,x) F(\varphi(t,x)).
\end{align}
\end{theorem}

@@

Simulating these processes and exploring parameter space,
one sooner or later comes upon a strange observation:
with certain parameter combinations, the population spontaneously forms a regular grid
of stable, more or less discrete patches, separated by areas with nearly no individuals,
as shown in Figure~\ref{fig:clumping}.
The phenomenon is discussed in Section~16.10 of \citet{haller2022evolutionary},
and has been described in similar models, e.g.,
by \citet{britton1990spatial,sasaki1997clumped,hernandezgarcia2004clustering,young2001reproductive}, and \citet{berestycki2009nonlocal}.
\comment{FROM ALISON: I rewrote the rest of this section - please check!}
For example, if the density-dependent effects of individuals extend farther (but not too much farther)
than the typical dispersal distance,
then depending on the interaction kernel
new offspring landing between two clumps
can effectively find themselves in competition with 
\emph{both} neighbouring clumps,
while individuals within a clump compete with only one.

@@

We now look at the second order term in~(\ref{generator prelimit}).
An individual at location $x$ gives birth 
to a surviving offspring at $y$ at rate
$$
\gamma(x,\eta) r(y,\eta) q_{\theta}(x, dy) ,
$$
and since this increments $\langle f, \eta \rangle$ by $f(y) / N$,
the contribution to the quadratic variation from birth events,
which occur at rate $\theta$ per individual 
(so, rate $N\theta |\eta|$ overall), is
$$
\int
    N \theta \gamma(x,\eta)
    \int \frac{1}{N^2} f^2(y) r(y,\eta)
    q_\theta(x,dy) 
\eta(dx) .
$$
Similarly, the increment in $\langle f, \eta\rangle$ resulting from 
the death of an individual at $x$ is $-f(x)/N$, and so combining with the 
above, the second order term in the generator takes the form
\begin{align*}
& G''(\langle f,\eta\rangle)
\frac{1}{2} N \theta
\left\{
    \int
        \gamma(x,\eta)
        \int \frac{1}{N^2}f^2(y)r(y,\eta)q_\theta(x,dy) 
    \eta(dx)
    +
    \int
        \mu_\theta(x,\eta)\frac{1}{N^2}f^2(x) 
    \eta(dx)
\right\} \\
&\qquad
= \frac{1}{2} G''(\langle f, \eta \rangle)
    \frac{\theta}{N}
	\int\left\{
		\gamma(x, \eta) \int f^2(y) r(y, \eta) q_\theta(x, dy) + f^2(x) \mu_\theta(x, \eta)
    \right\} \eta(dx) .
\end{align*}
\comment{FROM ALISON integral against eta in last line corrected}
Since $\int f^2(y) r(y, \eta) q_\theta(x, dy) \to f^2(x) r(x, \eta)$
and $r\gamma  + \mu_\theta = 2 r \gamma  - F / \theta \to 2 r \gamma$
as $\theta \to \infty$,
this converges to
\begin{align*}
\frac{\alpha}{2} G''(\langle f, \eta \rangle)
    \big\langle
	2 r(x, \eta) \gamma(x, \eta),
        \eta(dx)
    \big\rangle .
\end{align*}

@@

Putting these together,
the level $u$ of an individual at $x$ evolves according to:
\begin{equation} \label{eqn:dot_u}
    \dot u
    =
    - \theta\frac{u}{N} \left(N - u\right)
    \gamma(x, \eta) \int_{\IR^d} r(y, \eta) q_\theta(x, dy) 
    +
    \theta\mu_\theta(x,\eta) u .
\end{equation}
We shall write 
\begin{align*}
    b_\theta(x, \eta)
    :=
    \theta\left(
    \gamma(x,\eta) \int_{\IR^d} r(y, \eta) q_\theta(x, dy)
    -
    \mu_\theta(x,\eta)
    \right) ,
\end{align*}
which captures the local net difference between reproduction and death,
and
\begin{align} \label{c_defn}
    c_\theta(x, \eta)
    :=
    \frac{\theta}{N} \gamma(x, \eta) \int_{\IR^d} r(y, \eta) q_\theta(x, dy) ,
\end{align}
which captures the local rate of production of successful offspring.
\comment{FROM ALISON This sort of true because it corresponds to lookdown events, but when I read this again I was confused. Is there a better way to say it?} 
Recall from equation~\eqref{eqn:mu_defn} that
$F(x,\eta) = \theta(r(x,\eta)\gamma(x,\eta) - \mu_\theta(x,\eta))$,
and so 
\begin{equation}  \label{b_defn}
b_\theta(x, \eta)
=    \theta \gamma(x, \eta) \int_{\IR^d} \left( r(y, \eta) - r(x, \eta) \right) 
	q_\theta(x, dy)
    + F(x, \eta). 
\end{equation}
Under Assumptions~\ref{def:model_setup}, as $\theta \to \infty$,
$c_\theta(x, \eta)$ will tend to $\alpha \gamma(x, \eta) r(x, \eta)$
\begin{equation}
\label{eqn:b_limit}
    b_\theta(x, \eta) \to \gamma(x, \eta) \DG r(x, \eta) + F(x, \eta) .
\end{equation}

@@

\begin{definition}[Martingale Problem Characterisation]
    \label{defn:lookdown_mgale}
For given positive values of $N$ and $\theta$, define the generator $A^N$
by
\begin{equation}
	\label{eqn:lookdown_generator}
\begin{split}
& A^{N}f(\lp ) \\
&\quad =
    f(\lp)
    \,\sum_{(x,u)\in \lp}\,
    2 \gamma(x, \eta)
    \Bigg\{ \frac{1}{2N} \int_u^N g(x,u_1) du_1
            \frac{
                \theta \int_{\IR^d}
                (g(y,u) - g(x,u))
                r(y, \eta) q_{\theta}(x,dy)
            }{ g(x,u) }
        \\
    &\qquad\qquad\qquad\qquad\qquad\qquad\qquad {} +
        \frac{\theta}{N} \int_u^N
        \int_{\IR^d}\left(
            \frac{ g(y,u_1) + g(x,u_1) }{ 2 } - 1
        \right)
        r(y, \eta) q_{\theta}(x,dy)
        du_1
    \Bigg\}\\
    &\qquad\qquad {} +
    f(\lp) \sum_{(x,u)\in\lp}\,
    \left(
        c_\theta(x,\eta) u^2 - b_{\theta}(x,\eta)u
    \right)
    \frac{\partial_u g(x,u)}{g(x,u)}
    ,
\end{split}
\end{equation}
where $f(\lp) = \prod_{(x, u) \in \lp} g(x, u)$ is as defined in \eqref{eqn:test_functions},
and $\eta(\cdot) = \lp(\cdot \times [0, N]) / N$
as before. 
Given $\lp_0 \in {\mathcal C}_N$, 
we say that a ${\mathcal D}_{[0,\infty)}({\mathcal C}_N)$-valued process 
	\comment{FROM ALISON: we have not been consistent in the layout of 
	the notation for spaces of cadlag paths - it is already different in 
	the definition below!}
$(\xi_t)_{t \ge 0}$
   is a solution to the $(A^N, \lp_0)$ martingale problem
    if 
$f(\lp_t) - f(\lp_0)-\int_0^t A^N f(\lp_s) ds$ is a martingale (with 
respect to the natural filtration) for all 
test functions $f$ 
    as defined above.
\end{definition}

@@

\begin{definition}[Martingale Problem Characterisation, scaling limit]
    \label{defn:limiting_lookdown_mgale}
    Fix $\alpha \in [0, \infty)$, and define  
    test functions $f$ by 
    $f(\xi) = \prod_{(x, u) \in \xi} g(x, u)$ 
	with $g$ differentiable in $u$, smooth in $x$, satisfying 
$0\leq g(x,u)\leq 1$ and
    such that there exists a $u_0$ with $g(x, u) = 1$ for all $u > u_0$.
    Then, define the operator $A$ on such test functions by 
    \begin{align} \begin{split} \label{eqn:limiting_lookdown_generator}
    A f(\lp)
    &=
        f(\lp)  \sum_{(x, u) \in \xi}
        \gamma(x, \eta)
            \frac{
                \DG(g(\cdot, u) r(\cdot, \eta))(x) - g(x,u) \DG r(x,\eta)
            }{
                g(x, u)
            }
    \\ &\qquad {} +
        f(\lp) \sum_{(x, u) \in \xi}
        2 \alpha \gamma(x, \eta) r(x, \eta) \int_u^\infty (g(x, u_1) - 1) du_1
    \\ &\qquad {} +
        f(\lp) \sum_{(x, u) \in \xi}
        \left(
            \alpha \gamma(x, \eta) r(x, \eta) u^2
            -
            \left\{
                \gamma(x, \eta) \DG r(x, \eta) + F(x, \eta)
            \right\} u
        \right)
        \frac{\partial_u g(x, u)}{ g(x,u) }  .
    \end{split} \end{align}
    We say that a ${\cal D}([0,\infty), \mathcal{C})$-valued process $(\xi_t)_{t \ge 0}$
    is a solution to the $(A, \lp_0)$ martingale problem
    if it has initial distribution $\lp_0$
    and $f(\lp_t) - f(\lp_0)-\int_0^t Af(\lp_s) ds$ is a martingale (with
respect to the natural filtration) for all 
test functions $f$
    as defined above.
\end{definition}

@@

To make this precise,
define the spatial projection maps
$\kappa^N: \mathcal{M}(\IR^d \times [0, N]) \to \mathcal{M}(\IR^d)$
by $\kappa^N(\xi^N)(\cdot) = \xi^N(\cdot \times [0, N]) / N$,
and $\kappa: \mathcal{M}(\IR^d \times [0, \infty)) \to \mathcal{M}(\IR^d)$
by $\kappa(\xi)(\cdot) = \lim_{u_0 \to \infty} \xi(\cdot \times [0, u_0]) / u_0$.
We will also need an inverse notion:
for a measure $\lp^N$ on $\IR^d \times [0, N]$ and a $\sigma$-field $\mathcal{F}$,
we say that \emph{$\lp^N$ is conditionally uniform given $\mathcal{F}$}
if for all compactly supported $f$,
\begin{align} \label{eqn:conditionally_uniform}
    \IE[ e^{-\langle f, \lp \rangle} \;|\; \mathcal{F} ]
    &=
    e^{-\langle H_f^N, \kappa^N(\lp) \rangle} ,
\end{align}
where
$$ H_f^N(x) = - N \log \frac{1}{N} \int_0^N e^{-f(x, u)} du . $$
In other words, the $[0, N]$ components of $\lp$
are independent, uniformly distributed on $[0, N]$, and independent of 
$\kappa^N(\lp)$.
\comment{ALISON: added N, also made sigma $\sigma$ above}
Similarly, for a measure $\lp$ on $\IR^d \times [0, \infty)$
we say that \emph{$\lp$ is a conditionally Poisson random measure given $\mathcal{F}$}
if for all compactly supported $f$,
\begin{align} \label{eqn:conditionally_poisson}
    \IE[ e^{-\langle f, \lp \rangle} \;|\; \mathcal{F} ]
    &=
    e^{-\langle \int_0^\infty (1 - e^{-f(x, u)}) du, \kappa(\lp)(dx) \rangle} .
\end{align}
In other words, $\lp$ is conditionally Poisson with Cox measure $\kappa(\lp) \times \lambda$,
where $\lambda$ is Lesbegue measure.

\begin{proposition}
    \label{thm:mmt_application}
    If $\tilde \eta^N$ is a solution
    of the martingale problem of Definition~\ref{defn:mgale_construction}
    with initial distribution $\eta^N_0$
    then there exists a solution $\lp^N$
    of the $(A^N, \lp^N_0)$-martingale problem of Definition~\ref{defn:lookdown_mgale}
    such that $\eta^N = \kappa^N \circ \lp^N$
    has the same distribution on $D_{\measures}[0, \infty)$
    as $\tilde \eta^N$.
    Furthermore, for each $t$, $\lp^N_t$ is conditionally uniform 
	\comment{ALISON deleted excess given}
	given $\mathcal{F}^{\eta^N}_t$
    in the sense of \eqref{eqn:conditionally_uniform}.

@@

In this section we present formal proofs of the first two of our
three scaling limits.
%following sections we present formal proofs of our three scaling limits.
In Subsection~\ref{sec:population_density_proof} 
we prove Theorem~\ref{thm:nonlocal_convergence}, to obtain (both
stochastic and deterministic) limits in
which interactions between individuals in the population are nonlocal.
In Subsection~\ref{subsec:nonlocal to local}
we show how, in two important examples in which the nonlocal
limit is respectively a deterministic solution to a non-local equation of 
reaction-diffusion type and a deterministic solution to a nonlocal 
porous medium equation with an additional logistic growth term, 
one can pass to a further 
limit to obtain a classical PDE.
%In subsequent sections (Sections~\ref{subsec:one step convergence proof} and~\ref{sec:lookdown_proofs})
%we prove our third scaling limit, in which 
%we pass from our population model 
%directly to a reaction-diffusion equation without going via a nonlocal limit,
%and finally our results concerning both convergence of the lookdown representation
%and characterization of the lineage process in the deterministic case.
\comment{FROM ALISON: slightly reworded - deleted bit about onestep convergence
on the basis that we already point to that in the introduction}

>>>>

\comment{TODO: signpost where proofs can be found}

@@

\paragraph{Scaling limits with nonlocal interactions:}
Recall that the process $(\eta_t^N)_{t\geq 0}$ takes its values in the space
${\mathcal D}_{[0,\infty)}(\measures)$ of c\`adl\`ag
paths on $\measures$. We endow $\measures$ with the topology of weak convergence
and 
${\mathcal D}_{[0,\infty)}(\measures)$ with the Skorohod topology.
A sequence of processes taking values in 
${\mathcal D}_{[0,\infty)}(\measures)$ is said to be tight if the corresponding
sequence of distributions is tight,
i.e., that any infinite subsequence has a weakly convergent subsubsequence.
Our first main result establishes tightness of our rescaled population processes in 
the case in which interactions remain nonlocal under the scaling, and 
characterises limit points as solutions to a martingale problem.

@@

\begin{theorem}[Convergence to a PDE]
    \label{thm:local_convergence}
    Let $(\eta^N_t)_{t \geq 0}$
    be as defined in Definition~\ref{defn:mgale_construction} with
$r(x,m)\equiv 1\equiv \gamma(r,m)$, $F(x,m)\equiv F(m)$, 
$\rho^\epsilon_F$ a symmetric Gaussian density with variance
parameter $\epsilon^2$,
and $\DG=\Delta$. 
Further suppose that $F(m)$ is a polynomial with $F(m)\ind_{m\geq 0}$ bounded above.
Assume that $\langle 1,\eta_0\rangle$ is bounded, and that for all 
$x\in\IR^d$ and $k\in \IN$,
	\[
		\limsup_{\epsilon\to 0} 
		\IE\big[\rho_F^\epsilon*\eta_0(x)^k\big]<\infty,
	\]
	and 
\[
    \limsup_{\epsilon \to 0} \int \IE\big[\rho_F^\epsilon*\eta_0(x)^k\big] dx<\infty.
\]
Finally assume that $N \to \infty$, $\theta \to \infty$
and $\epsilon \to 0$ in such a way that
\begin{equation}
\frac{1}{\theta\epsilon^2}+\frac{\theta}{N\epsilon^d}\to 0.
\end{equation}
Then the sequence of 
    ${\mathcal D}_{[0,\infty)}(\measures)$-valued
stochastic processes $\big(\rho_F^\epsilon*\eta^N_t(x)dx\big)_{t \ge 0}$
converges weakly
to a measure-valued process with a density $\varphi(t, x)$
that solves
\begin{align}
	\label{target equation in local convergence}
        \partial_t \varphi(t, x)=\Delta \varphi (t,x) +\varphi(t,x) F(\varphi(t,x)).
\end{align}
\end{theorem}

@@

Simulating these processes and exploring parameter space,
one sooner or later comes upon a strange observation:
with certain parameter combinations, the population spontaneously forms a regular grid
of stable, more or less discrete patches, separated by areas with nearly no individuals,
as shown in Figure~\ref{fig:clumping}.
The phenomenon is discussed in Section~16.10 of \citet{haller2022evolutionary},
and has been described in similar models, e.g.,
by \citet{britton1990spatial,sasaki1997clumped,hernandezgarcia2004clustering,young2001reproductive}, and \citet{berestycki2009nonlocal}.
For example, if the density-dependent effects of individuals extend farther (but not too much farther)
than the typical dispersal distance,
then depending on the interaction kernel
new offspring landing between two clumps
can effectively find themselves in competition with 
\emph{both} neighbouring clumps,
while individuals within a clump compete with only one.

@@

We now look at the second order term in~(\ref{generator prelimit}).
An individual at location $x$ gives birth 
to a surviving offspring at $y$ at rate
$$
\gamma(x,\eta) r(y,\eta) q_{\theta}(x, dy) ,
$$
and since this increments $\langle f, \eta \rangle$ by $f(y) / N$,
the contribution to the quadratic variation from birth events,
which occur at rate $\theta$ per individual 
(so, rate $N\theta |\eta|$ overall), is
$$
\int
    N \theta \gamma(x,\eta)
    \int \frac{1}{N^2} f^2(y) r(y,\eta)
    q_\theta(x,dy) 
\eta(dx) .
$$
Similarly, the increment in $\langle f, \eta\rangle$ resulting from 
the death of an individual at $x$ is $-f(x)/N$, and so combining with the 
above, the second order term in the generator takes the form
\begin{align*}
& G''(\langle f,\eta\rangle)
\frac{1}{2} N \theta
\left\{
    \int
        \gamma(x,\eta)
        \int \frac{1}{N^2}f^2(y)r(y,\eta)q_\theta(x,dy) 
    \eta(dx)
    +
    \int
        \mu_\theta(x,\eta)\frac{1}{N^2}f^2(x) 
    \eta(dx)
\right\} \\
&\qquad
= \frac{1}{2} G''(\langle f, \eta \rangle)
    \frac{\theta}{N}
	\int\left\{
		\gamma(x, \eta) \int f^2(y) r(y, \eta) q_\theta(x, dy) + f^2(x) \mu_\theta(x, \eta)
    \right\} \eta(dx) .
\end{align*}
Since $\int f^2(y) r(y, \eta) q_\theta(x, dy) \to f^2(x) r(x, \eta)$
and $r\gamma  + \mu_\theta = 2 r \gamma  - F / \theta \to 2 r \gamma$
as $\theta \to \infty$,
this converges to
\begin{align*}
\frac{\alpha}{2} G''(\langle f, \eta \rangle)
    \big\langle
	2 r(x, \eta) \gamma(x, \eta),
        \eta(dx)
    \big\rangle .
\end{align*}

@@

Putting these together,
the level $u$ of an individual at $x$ evolves according to:
\begin{equation} \label{eqn:dot_u}
    \dot u
    =
    - \theta\frac{u}{N} \left(N - u\right)
    \gamma(x, \eta) \int_{\IR^d} r(y, \eta) q_\theta(x, dy) 
    +
    \theta\mu_\theta(x,\eta) u .
\end{equation}
We shall write 
\begin{align*}
    b_\theta(x, \eta)
    :=
    \theta\left(
    \gamma(x,\eta) \int_{\IR^d} r(y, \eta) q_\theta(x, dy)
    -
    \mu_\theta(x,\eta)
    \right) ,
\end{align*}
which captures the local net difference between reproduction and death,
and
\begin{align} \label{c_defn}
    c_\theta(x, \eta)
    :=
    \frac{\theta}{N} \gamma(x, \eta) \int_{\IR^d} r(y, \eta) q_\theta(x, dy) ,
\end{align}
which captures the local rate of production of successful offspring.
\comment{FROM ALISON This sort of true because it corresponds to lookdown events, but when I read this again I was confused. Is there a better way to say it?} 
\comment{PETER: I'm not sure which bit is confusing?}
Recall from equation~\eqref{eqn:mu_defn} that
$F(x,\eta) = \theta(r(x,\eta)\gamma(x,\eta) - \mu_\theta(x,\eta))$,
and so 
\begin{equation}  \label{b_defn}
b_\theta(x, \eta)
=    \theta \gamma(x, \eta) \int_{\IR^d} \left( r(y, \eta) - r(x, \eta) \right) 
	q_\theta(x, dy)
    + F(x, \eta). 
\end{equation}
Under Assumptions~\ref{def:model_setup}, as $\theta \to \infty$,
$c_\theta(x, \eta)$ will tend to $\alpha \gamma(x, \eta) r(x, \eta)$
\begin{equation}
\label{eqn:b_limit}
    b_\theta(x, \eta) \to \gamma(x, \eta) \DG r(x, \eta) + F(x, \eta) .
\end{equation}

@@

\begin{definition}[Martingale Problem Characterisation]
    \label{defn:lookdown_mgale}
For given positive values of $N$ and $\theta$, define the generator $A^N$
by
\begin{equation}
	\label{eqn:lookdown_generator}
\begin{split}
& A^{N}f(\lp ) \\
&\quad =
    f(\lp)
    \,\sum_{(x,u)\in \lp}\,
    2 \gamma(x, \eta)
    \Bigg\{ \frac{1}{2N} \int_u^N g(x,u_1) du_1
            \frac{
                \theta \int_{\IR^d}
                (g(y,u) - g(x,u))
                r(y, \eta) q_{\theta}(x,dy)
            }{ g(x,u) }
        \\
    &\qquad\qquad\qquad\qquad\qquad\qquad\qquad {} +
        \frac{\theta}{N} \int_u^N
        \int_{\IR^d}\left(
            \frac{ g(y,u_1) + g(x,u_1) }{ 2 } - 1
        \right)
        r(y, \eta) q_{\theta}(x,dy)
        du_1
    \Bigg\}\\
    &\qquad\qquad {} +
    f(\lp) \sum_{(x,u)\in\lp}\,
    \left(
        c_\theta(x,\eta) u^2 - b_{\theta}(x,\eta)u
    \right)
    \frac{\partial_u g(x,u)}{g(x,u)}
    ,
\end{split}
\end{equation}
where $f(\lp) = \prod_{(x, u) \in \lp} g(x, u)$ is as defined in \eqref{eqn:test_functions},
and $\eta(\cdot) = \lp(\cdot \times [0, N]) / N$
as before. 
Given $\lp_0 \in {\mathcal C}_N$, 
we say that a ${\mathcal D}_{[0,\infty)}({\mathcal C}_N)$-valued process 
$(\xi_t)_{t \ge 0}$ is a solution to the $(A^N, \lp_0)$ martingale problem
if $f(\lp_t) - f(\lp_0)-\int_0^t A^N f(\lp_s) ds$ is a martingale (with 
respect to the natural filtration) for all 
test functions $f$ as defined above.
\end{definition}

@@

\begin{definition}[Martingale Problem Characterisation, scaling limit]
    \label{defn:limiting_lookdown_mgale}
    Fix $\alpha \in [0, \infty)$, and define  
    test functions $f$ by 
    $f(\xi) = \prod_{(x, u) \in \xi} g(x, u)$ 
	with $g$ differentiable in $u$, smooth in $x$, satisfying 
$0\leq g(x,u)\leq 1$ and
    such that there exists a $u_0$ with $g(x, u) = 1$ for all $u > u_0$.
    Then, define the operator $A$ on such test functions by 
    \begin{align} \begin{split} \label{eqn:limiting_lookdown_generator}
    A f(\lp)
    &=
        f(\lp)  \sum_{(x, u) \in \xi}
        \gamma(x, \eta)
            \frac{
                \DG(g(\cdot, u) r(\cdot, \eta))(x) - g(x,u) \DG r(x,\eta)
            }{
                g(x, u)
            }
    \\ &\qquad {} +
        f(\lp) \sum_{(x, u) \in \xi}
        2 \alpha \gamma(x, \eta) r(x, \eta) \int_u^\infty (g(x, u_1) - 1) du_1
    \\ &\qquad {} +
        f(\lp) \sum_{(x, u) \in \xi}
        \left(
            \alpha \gamma(x, \eta) r(x, \eta) u^2
            -
            \left\{
                \gamma(x, \eta) \DG r(x, \eta) + F(x, \eta)
            \right\} u
        \right)
        \frac{\partial_u g(x, u)}{ g(x,u) }  .
    \end{split} \end{align}
    We say that a ${\mathcal D}_{[0,\infty)}(\mathcal{C})$-valued process $(\xi_t)_{t \ge 0}$
    is a solution to the $(A, \lp_0)$ martingale problem
    if it has initial distribution $\lp_0$
    and $f(\lp_t) - f(\lp_0)-\int_0^t Af(\lp_s) ds$ is a martingale (with
respect to the natural filtration) for all 
test functions $f$
    as defined above.
\end{definition}

@@

To make this precise,
define the spatial projection maps
$\kappa^N: \mathcal{M}(\IR^d \times [0, N]) \to \mathcal{M}(\IR^d)$
by $\kappa^N(\xi^N)(\cdot) = \xi^N(\cdot \times [0, N]) / N$,
and $\kappa: \mathcal{M}(\IR^d \times [0, \infty)) \to \mathcal{M}(\IR^d)$
by $\kappa(\xi)(\cdot) = \lim_{u_0 \to \infty} \xi(\cdot \times [0, u_0]) / u_0$.
We will also need an inverse notion:
for a measure $\lp^N$ on $\IR^d \times [0, N]$ and a $\sigma$-field $\mathcal{F}$,
we say that \emph{$\lp^N$ is conditionally uniform given $\mathcal{F}$}
if for all compactly supported $f$,
\begin{align} \label{eqn:conditionally_uniform}
    \IE[ e^{-\langle f, \lp \rangle} \;|\; \mathcal{F} ]
    &=
    e^{-\langle H_f^N, \kappa^N(\lp) \rangle} ,
\end{align}
where
$$ H_f^N(x) = - N \log \frac{1}{N} \int_0^N e^{-f(x, u)} du . $$
In other words, the $[0, N]$ components of $\lp$
are independent, uniformly distributed on $[0, N]$, and independent of 
$\kappa^N(\lp)$.
Similarly, for a measure $\lp$ on $\IR^d \times [0, \infty)$
we say that \emph{$\lp$ is a conditionally Poisson random measure given $\mathcal{F}$}
if for all compactly supported $f$,
\begin{align} \label{eqn:conditionally_poisson}
    \IE[ e^{-\langle f, \lp \rangle} \;|\; \mathcal{F} ]
    &=
    e^{-\langle \int_0^\infty (1 - e^{-f(x, u)}) du, \kappa(\lp)(dx) \rangle} .
\end{align}
In other words, $\lp$ is conditionally Poisson with Cox measure $\kappa(\lp) \times \lambda$,
where $\lambda$ is Lesbegue measure.

\begin{proposition}
    \label{thm:mmt_application}
    If $\tilde \eta^N$ is a solution
    of the martingale problem of Definition~\ref{defn:mgale_construction}
    with initial distribution $\eta^N_0$
    then there exists a solution $\lp^N$
    of the $(A^N, \lp^N_0)$-martingale problem of Definition~\ref{defn:lookdown_mgale}
    such that $\eta^N = \kappa^N \circ \lp^N$
    has the same distribution on $D_{\measures}[0, \infty)$
    as $\tilde \eta^N$.
    Furthermore, for each $t$, $\lp^N_t$ is conditionally uniform 
	given $\mathcal{F}^{\eta^N}_t$
    in the sense of \eqref{eqn:conditionally_uniform}.

@@

In this section we present formal proofs of the first two of our
three scaling limits.
%following sections we present formal proofs of our three scaling limits.
In Subsection~\ref{sec:population_density_proof} 
we prove Theorem~\ref{thm:nonlocal_convergence}, to obtain (both
stochastic and deterministic) limits in
which interactions between individuals in the population are nonlocal.
In Subsection~\ref{subsec:nonlocal to local}
we show how, in two important examples in which the nonlocal
limit is respectively a deterministic solution to a non-local equation of 
reaction-diffusion type and a deterministic solution to a nonlocal 
porous medium equation with an additional logistic growth term, 
one can pass to a further 
limit to obtain a classical PDE.
%In subsequent sections (Sections~\ref{subsec:one step convergence proof} and~\ref{sec:lookdown_proofs})
%we prove our third scaling limit, in which 
%we pass from our population model 
%directly to a reaction-diffusion equation without going via a nonlocal limit,
%and finally our results concerning both convergence of the lookdown representation
%and characterization of the lineage process in the deterministic case.
